Graph Neural Networks For Multi-Image Matching

### 2019-09-25 latex/iclr2020/main.tex
Added Sid (and a few of Daphne's) comments

\begin{abstract}
    Image feature matching is a fundamental part of many geometric computer vision applications, and using multiple images can improve performance.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence is difficult or expensive to acquire.
    In addition, geometric consistency losses can be added at training time, even if the information is not available in the test set, unlike previous approaches which optimize cycle consistency directly.
    To the best of our knowledge, no other works have used graph neural networks for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D image locations from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching (\cite{lowe2004distinctive}) followed by RANSAC (\cite{fischler1981random}) have been the standard for decades.
However RANSAC-based approaches are limited to matching pairs of images, which can lead to global inconsistencies in the matching.
% TODO: Mention trifocal tensors
Other works, such as Wang et al.~\cite{wang2017multi}, have shown improvement in performance by optimizing cycle consistency, i.e. enforcing the pairwise feature matches to be globally consistent.

However, these multi-view consistency algorithms struggle in distributed and robust settings.
Having image features suited for this task would help improve performance, and deep learning has revolutionized how image features are computed (\cite{yi2016lift}).
In this paper, we want to leverage the power of deep representations in order to compute feature descriptors that are robust across multiple views.
% Typically when applied to image or feature matching, deep neural networks (DNNs) are trained with photometric losses, directly using pixel color values.
% However direct, pixel-based losses have are not robust to the many confounding factors in the environment, such as lighting and view-point changes.
% Using features designed to deal with these confounding factors improves robustness to them, and learning may improve robustness even further.

% \begin{figure}[t]
% \begin{center}
%   % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
%   \includegraphics[width=0.8\linewidth]{figures-CycleConsistencyBasic-v2.pdf}
% \end{center}
%   \caption{
%     Example of multi-image matching and cycle consistency, with images from Rome16K dataset \cite{li2010location}.
%     Pairwise matches between image can lead to global inconsistencies in the matching.
%     In this example, an error in matching leads to an inconsistent cycle (shown in red), ideally these pairwise matches would be consistent.
%     We improve the quality of pairwise matches in the multi-view matching problem by training a neural net with cycle consistency.
%   }
% \label{fig:cycconsistex}
% \end{figure}

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures-CycleConsistencyMainFigure-v4.pdf}
\end{center}
  \caption{
    An illustration of the approach of this work.
    The Graph Neural Neural Network (GNN) (\cite{battaglia2018relational}) takes as input the graph of matches and then outputs a low rank embedding of the adjacency matrix of the graph.
    The GNN operates on an embedding over the vertices of the graph.
    In the figure, the GNN vertex embeddings are represented by different colors.
    The final embedding is used to construct a pairwise similarity matrix, which we train to be a low dimensional cycle-consistent representation of the graph adjacency matrix, thus pruning the erroneous matches.
    We train the network using a reconstruction loss on the similarity matrix with the noisy adjacency matrix, and thus do not need ground truth matches.
    In addition, we can use geometric consistency information, such as epipolar constraints, to assist training the network.
  }
\label{fig:pipeline}
\end{figure*}

Unfortunately, there are obstacles to applying multi-view constraints directly to deep learning. 
To train networks, we need a large amount of labeled data.
In the case of multi-image feature matching, one would need hand labeled point correspondences between images, which are difficult and expensive to obtain.
Multi-view constraints are formulated in terms of sparse features, which traditional convolutional neural nets are not designed to handle.
Additionally, in the case of multi-image feature matching, geometric constraints would be helpful in rejecting outlier matches.
Geometric constraints, such as the epipolar constraint, can help disambiguate between visually similar features during training, and help the network learn better features robust to this kind of noise.

In this work, we propose to solve these problems using Graph Neural Networks (GNNs) to operate on the correspondence graph.
The proposed method works directly on the correspondence graph, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss - the cycle consistency loss - to train the network, and thus avoiding the difficulty of expensive hand labeling.
Geometric consistency losses can aid training, even if such information is not available at test time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings, using a matrix factorization loss (\cite{zhou2015multi, leonardos2016distributed}).
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster (to achieve comparable accuracy) than methods which must solve an optimization problem every time.
We perform experiments on the Rome16K dataset (\cite{li2010location}) to test the effectiveness of our method compared to optimization based methods.
Our contributions in this work are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\end{itemize}

@@

\subsection{Feature Matching}
Matching has a rich history of research in computer vision.
Much work has be done using hand-crafted feature descriptors such as SIFT (\cite{lowe2004distinctive}), SURF (\cite{bay2006surf}), BRIEF (\cite{calonder2012brief}), or ORB (\cite{mur2015orb}).
RANSAC \cite{fischler1981random} is the most widely used robust estimation technique to filter out outliers from the matches.
The combination of RANSAC and hand-crafted feature descriptors has constituted the bulk of the matching literature for the last 40 years.
Finally, graph matching (\cite{suh2015subgraph, hu2016distributable}) can be used as a final step for more robust matches.

\subsection{Multi-image Matching}
Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section 3.3).
\cite{pachauri2013solving} and \cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
However, the low Gaussian noise assumption is not realistic.
\cite{zhou2015multi} and \cite{wang2017multi} use most sophisticated optimization techniques on the matching matrix and thus produce more robust solutions.
\cite{leonardos2016distributed} implement a distributed optimization scheme to solve for cycle consistency.
As an alternative to optimization based techniques, \cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.
Moving away from feature matching, \cite{zach2010disambiguating} uses cycle consistency like constraints on pose graphs quite effectively.
To the best of our knowledge, we are the first to use graph neural networks for multi-image matching.

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets such as \cite{zagoruyko2015learning}; \cite{yi2016lift}; and \cite{brachmann2017dsac}.
This approach is limited if one do not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as \cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, \cite{yi2018learning} attempts to improve correspondences by learning match probabilities for RANSAC for greater robustness and speed.
However, they only focus on two view matching and do not exploit the advantages of the correspondence structure.
Note that while \cite{zhu2017unpaired} use cycle consistency in their loss, their method is restricted to pairwise cycle consistency (i.e. enforcing consistency by going back and forth between two images).
We use multi-image cycle consistency, which require consistency between 3 or more images.

\subsection{Graph Neural Networks}
Graph neural networks have received more attention recently e.g. \cite{bronstein2017geometric, bruna2013spectral, defferrard2016convolutional, kipf2017semi, scarselli2009graph, gama2018mimo, gama2018convolutional}; and \cite{battaglia2018relational}.
Classical so-called Spectral methods used the eigenvectors of the graph Laplacian to compute convolutions as in \cite{bruna2013spectral}, but requires an a-priori known graph structure. 
Non-spectral methods do not require a-priori knowledge, as seen in \cite{bronstein2017geometric, kipf2017semi, scarselli2009graph}; and \cite{gama2018convolutional}.
Most of these methods use polynomials of the graph Laplacian to compute neighborhood averages.
\cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in \cite{battaglia2018relational}.

@@

\begin{figure}[t]
\begin{subfigure}[b]{.55\linewidth}
  \centering
  \includegraphics[width=0.95\linewidth]{figures-GeometricConsistency-v2.pdf}
  \caption{Illustrated here is an example of how the geometric loss is computed for one feature.}
  \label{fig:1b}
  \label{fig:geoconsist}
\end{subfigure}
\begin{subfigure}[b]{.30\textwidth}
  \centering
  % \includegraphics[width=0.65\linewidth]{figures-GeometricLoss.pdf}
  \includegraphics[height=180px]{figures-GeometricLoss.pdf}
  \caption{}
  \label{fig:3b}
  \label{fig:geomloss}
\end{subfigure}
\caption{
  \textbf{(a)} Errors are computed via absolute distance from the epipolar line, as expressed by (\ref{eq:essential_constraint}) via the epipolar constraint.
  The epipolar line is the line of projection of the feature on the first image, projected onto to the second.
  The distance to this line on the second image indicates how likely that point is to correspond geometrically to the original feature.
  There can be false positives along the projected line, as shown by the square feature in the figure, but other points will be eliminated, such as the hexagonal feature.
  \textbf{(b)} Training curves with and without Geometric Training loss, described in \ref{eq:geom_cost2}.
  The geometric training loss improves testing performance.
  While the training is higher due to the additional loss terms, the ground truth L1 error is substantially better with the Geometric Loss.
  Best viewed in color.
}
\label{fig:1}
\end{figure}

\begin{figure}[t]
\begin{subfigure}[b]{.45\linewidth}
  \centering
  \includegraphics[width=0.95\linewidth]{figures-UniverseOfFeatures-v2.pdf}
  \caption{Illustration of the universe of features.}
  \label{fig:1a}
  \label{fig:universefeatures}
\end{subfigure}%
\begin{subfigure}[b]{.55\linewidth}
  \centering
  \includegraphics[width=0.75\linewidth]{figures-EmbeddingsViz.png}
  % \includegraphics[height=150px]{figures-EmbeddingsViz.png}
  \caption{}
  \label{fig:1b}
  \label{fig:embeddingsviz}
\end{subfigure}
\caption{
  \textbf{(a)} Each feature in each image corresponds to a 3D point in the scene.
  We can construct cycle consistent embeddings of the features by mapping each one to the one-hot vector of its corresponding 3D point.
  While there can be many features, there are fewer 3D points and thus this corresponds to a low rank factorization of the correspondence matrix.
  \textbf{(b)} Visualization of the learned embeddings.
  On the left we have the raw outputs, which are difficult to interpret.
  In the center, we rotated the features to best match the ground truth for a more interpretable visualization.
  On the right, we have the ground truth embeddings, given as indicator vectors for which feature in the world the points correspond to.
  Best viewed in color.
}
\label{fig:1}
\end{figure}

Our goal is to learn optimal features that capture multiple image views by filtering out noisy feature matches.
An outline of our approach can be seen in figure \ref{fig:pipeline}.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
We do this by training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Graphs $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ have a set of vertices $\mathcal{V}$ and of edges $\mathcal{E} \subseteq \mathcal{V} \times \mathcal{V}$ and implicitly in the subsequent discussions we assume the edges are directed.
For a vertex $v \in \mathcal{V}$ we use $\mathcal{N}^{h}(v)$ to denote the $h$-hop neighbors of $v$, with the superscript left out for 1-hop neighbors.
Similarly $\mathcal{E}(v)$ is used to denote the edges associated with $v$.
To denote the vertices connected to an edge $e \in \mathcal{E}$ we write $e(v_1, v_2)$.
\subsection{Correspondence Graph}
We assume there is an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, with an associated adjacency matrix $\mA$.
The graph is constructed from putative correspondences of image features across images, typically constructed using feature descriptor distance (e.g. SIFT feature distance).
While there are many interesting methods for computing these putative correspondences (\cite{suh2015subgraph, yi2018learning}), we do not explore them in this work.
Typically putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
The ambiguity in the matches could come from repeated structures in the scene, insufficiently informative low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

@@

\subsection{Graph Neural Networks}
As input to our method we are given a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with the features described last section $\vf_v \; \forall v \in \mathcal{V}$ and $\vf_e \; \forall e \in \mathcal{E}$.
As with any neural network, GNNs have layered outputs.
We describe the output of layer $k$ as $\vf_v^{(k)} \in \sR^{m_k} \, \forall v \in \mathcal{V}$ and $\vf_e^{(k)} \in \sR^{p_k}\, \forall e \in \mathcal{E}$, with the initial embeddings denoted $\vf_v^{(0)} = \vf_v$ and $\vf_e^{(0)} = \vf_e$.
It will be useful to write these features as matrices in later sections, so we denote the vertex embedding matrix as $\mF_V^{(k)}$ and the edge embedding matrix as $\mF_E^{(k)}$.
If a superscript is not specified then it refers to the final output of the network.

First we describe older methods of GNNs to give context, then we describe the method we use in this work.
Many older methods assume we have the adjacency matrix $\mA$ of the graph known a-priori \cite{bruna2013spectral}, and can encode graph convolutions using the eigenvectors of $\mA$.
However, we do not have this luxury, as the correspondence structure changes from image set to image set, and thus we use non-spectral Graph Neural Networks.
There are many variations on non-spectral methods, often which ultimately amount to message passing between vertices of the graph with learned non-linear transformations between various steps (\cite{kipf2017semi, defferrard2016convolutional, gama2018mimo, gama2018convolutional}).
Some works such as \cite{gama2019convolutional} use pooling operations on the vertices to make the graph smaller and thus aid computation, but as we need labels on every vertex of the original graph, so we cannot use this.
Most of GNNs used in these works can be expressed mathematically as:
\begin{align*}
\tilde{\vf}_v^{(k+1)} =&\; \sigma\left( b^{(k)} + \mW_0^{k} \vf_{v}^{(k)} + \sum_{h=0}^H \sum_{v' \in \mathcal{N}^{h}(v)} f_{e(v,v')} \mW_{h}^{k} \vf_{v'}^{(k)} \right) \\
\end{align*}
The weights/biases $\mW_{h}^{k}$, $b^{(k)}$ are all learned, with no learning done on the edge weights $f_{e(v,v')}$. 
Note that this is just sums or averages over $h$-hop neighborhoods, where the weights on the edges remain static through the computation.
Given that we are trying to prune edges, it would be sensible to add features over edges to learn which ones to prune and which to keep such as in \cite{scarselli2009graph}.

Therefore, in this work we use the method and implementation described in \cite{battaglia2018relational} (more specifically repetitions of the architecture described in \cite{battaglia2016interaction}).
This work is similar to previous works with one key difference: edges also have weights.
Thus therefore there is intermediate processing on the edges before information is passed to the vertices.

Mathematically, this is expressed as: % TODO: Maybe make this MLP oriented?
\begin{align}
\tilde{\vf}_{e(v_1,v_2)}^{(k+1)} =&\; \sigma \left(a^{(k)} + \mU_1^{(k)} \vf_{e}^{(k)} + \mU_1^{(k)} \vf_{v_1}^{(k)} + \mU_2^{(k)} \vf_{v_2}^{(k)} \right) \\
\tilde{\vf}_{v}^{(k+1)} =&\; \sigma\left(b^{(k)} + \mW_0^{(k)} \vf_{v}^{(k)} + \sum_{e \in \mathcal{E}(v)} \mW_1^{(k)} \vf_{e}^{(k+1)} \right) 
\end{align}
Here the learned weights are denoted $\mW$ and $\mU$, and the biases $a^{(k)}$ and  $b^{(k)}$.
In \cite{battaglia2018relational}, they allow for more sophisticated aggregation functions, but in this work we simply use the mean function.
In practice, we use MLPs in message passes between vertices and edges for better expressiveness

\subsection{Cycle Consistency}

Let $\mM$ be the noiseless set of matches between our features, with $\mM_{ij}$ being the partial permutation representing the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then for all $i, j, k$:
\begin{equation}
\mM_{ij} = \mM_{ik} \mM_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency (\cite{zhou2015multi, wang2017multi, leonardos2016distributed}).
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a ground-truth matching matrix $\mX_i$.
Then the cycle consistency constraint becomes:
\begin{equation}
\mM_{ij} = \mX_{i}\mX_{j}^\top
\label{eq:cycconsist2}
\end{equation}

>>>>

\begin{abstract}
    In geometric computer vision applications, multi-image feature matching gives more accurate and robust solutions compared to simple two-image matching.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence can be difficult or expensive to acquire.
    % TODO: Ask Sid if this is any good
    Geometric consistency losses are added to aid training, though unlike optimization based methods no geometric information is necessary at inference time.
    To the best of our knowledge, no other works have used graph neural networks for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}
%     Image feature matching is a fundamental part of many geometric computer vision applications, and using multiple images can improve performance.
%     In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
%     We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence is difficult or expensive to acquire.
%     In addition, geometric consistency losses can be added at training time, even if the information is not available in the test set, unlike previous approaches which optimize cycle consistency directly.
%     To the best of our knowledge, no other works have used graph neural networks for multi-image feature matching.
%     Our experiments show that our method is competitive with other optimization based approaches.

@@

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D feature positions from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching \citep{lowe2004distinctive} followed by RANSAC \citep{fischler1981random} have been the standard for decades.
However RANSAC-based approaches are limited to matching pairs of images, which can lead to global inconsistencies in the matching.
% TODO: Mention trifocal tensors
Other works, such as \cite{wang2017multi}, have shown improvement in performance by optimizing cycle consistency, i.e. enforcing the pairwise feature matches to be globally consistent.

However, these multi-view consistency algorithms struggle in distributed and robust settings.
Having image features suited for this task would help improve performance, and deep learning has revolutionized how image features are computed \citep{yi2016lift}.
In this paper, we want to leverage the power of deep representations in order to compute feature descriptors that are robust across multiple views.

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures-CycleConsistencyMainFigure-v4.pdf}
\end{center}
  \caption{
    An illustration of the approach of this work.
    The Graph Neural Neural Network (GNN) \citep{battaglia2018relational} takes as input the graph of matches and then outputs a low rank embedding of the adjacency matrix of the graph.
    The GNN operates on an embedding over the vertices of the graph.
    In the figure, the GNN vertex embeddings are represented by different colors.
    The final embedding is used to construct a pairwise similarity matrix, which we train to be a low dimensional cycle-consistent representation of the graph adjacency matrix, thus pruning the erroneous matches.
    We train the network using a reconstruction loss on the similarity matrix with the noisy adjacency matrix, and thus do not need ground truth matches.
    In addition, we can use geometric consistency information, such as epipolar constraints, to assist training the network.
  }
\label{fig:pipeline}
\end{figure*}

Unfortunately, there are obstacles to applying multi-view constraints directly to deep learning. 
Multi-view constraints are formulated in terms of sparse features, which traditional convolutional neural nets are not designed to handle.
Thus we will need a new architecture to handle such constraints.
More fundamentally, deep neural networks need large amounts of labeled data to train.
In the case of multi-image feature matching, one would need hand-labeled point correspondences between images, which can be difficult and expensive to obtain.
Consequently unsupervised training is a more practical approach.
In the absence of direct supervision, the additional signal of geometric constraints can help disambiguate visually similar features and reject outliers.
Thus incorporating such constraints is important in training a network to solve this task.

In this work, we address these concerns using Graph Neural Networks (GNNs).
The proposed method works directly on the graph of correspondences between the image features, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss, the cycle consistency loss, to train the network, thus avoiding the difficulty of expensive hand-labeling.
We use geometric consistency losses to aid training, though no geometric information is used at inference time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings, using a matrix factorization loss \citep{zhou2015multi, leonardos2016distributed}.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster to achieve comparable accuracy than methods which must solve an optimization problem every time.
We perform experiments on the Rome16K dataset \citep{li2010location} to test the effectiveness of our method compared to optimization based methods.
Our contributions in this work are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\end{itemize}

@@

\subsection{Feature Matching}
Image feature matching has a rich history of research in computer vision.
Much work has be done using hand-crafted feature descriptors such as SIFT \citep{lowe2004distinctive}, SURF \citep{bay2006surf}, BRIEF \citep{calonder2012brief}, or ORB \citep{mur2015orb}.
RANSAC \cite{fischler1981random} is the most widely used robust estimation technique to filter out outliers from the matches.
The combination of RANSAC and hand-crafted feature descriptors has constituted the bulk of the matching literature for the last 40 years.
More recently \cite{suh2015subgraph} and \cite{hu2016distributable} have shown graph matching of the features can be added for more robust matches between images.

\subsection{Multi-image Matching}
Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section 3.3).
\cite{pachauri2013solving} and \cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
However, the assumption of low Gaussian noise is not realistic.
\cite{zhou2015multi} and \cite{wang2017multi} use more sophisticated optimization techniques on the matching matrix and thus produce more robust solutions.
\cite{leonardos2016distributed} implement a distributed optimization scheme to solve for cycle consistency.
As an alternative to optimization based techniques, \cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.
Moving away from feature matching, \cite{zach2010disambiguating} uses cycle consistency like constraints on pose graphs quite effectively.
To the best of our knowledge, we are the first to use graph neural networks for multi-image matching.

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets such as \cite{zagoruyko2015learning}; \cite{yi2016lift}; and \cite{brachmann2017dsac}.
This approach is limited if one do not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as \cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, \cite{yi2018learning} attempts to improve correspondences by learning match probabilities for RANSAC for greater robustness and speed.
However, they only focus on two view matching and do not exploit the advantages of the correspondence structure.
Like us \cite{zhu2017unpaired} use cycle consistency in their loss, however their method is restricted to pairwise cycle consistency (i.e. enforcing consistency by going back and forth between two images).
Our method can be applied to 3 or more images. 
\cite{hartmann2017learned} in fact learns multi-image matching but requires heavy supervision from 3D object reconstructions, which are not available in the context we are in.

\subsection{Graph Neural Networks}
Graph neural networks, true to their name, are deep neural networks operating in graph domains.
Multi-image matching is novel application of them, as they are more typically used for applications such as citation networks or recommendation systems.
They have received more attention recently \citep{bronstein2017geometric, defferrard2016convolutional, kipf2017semi, scarselli2009graph, gama2018mimo, gama2018convolutional, battaglia2018relational}.
The first methods to learn neural networks over graphs were the so-called Spectral methods.
They used the eigenvectors of the graph Laplacian to compute convolutions as in \cite{bruna2013spectral}, which can learn graph specific convolution kernels, but require an a-priori known graph structure. 
Newer non-spectral methods do not require a-priori knowledge, as seen in \cite{bronstein2017geometric, kipf2017semi, scarselli2009graph}; and \cite{gama2018convolutional}.
Most of these methods use polynomials of the graph Laplacian to compute neighborhood averages.
\cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in \cite{battaglia2018relational}.
The closest work on Graph Neural Networks to ours is \cite{kipf2016variational}, which uses GNNs to reconstruct the adjacency matrices, though not in a geometric context.

@@

\begin{figure}[t]
\begin{subfigure}[b]{.45\linewidth}
  \centering
  \includegraphics[width=0.95\linewidth]{figures-UniverseOfFeatures-v2.pdf}
  \caption{}
  \label{fig:1a}
  \label{fig:universefeatures}
\end{subfigure}%
\begin{subfigure}[b]{.55\linewidth}
  \centering
  \includegraphics[width=0.75\linewidth]{figures-EmbeddingsViz.png}
  % \includegraphics[height=150px]{figures-EmbeddingsViz.png}
  \caption{}
  \label{fig:1b}
  \label{fig:embeddingsviz}
\end{subfigure}
\caption{
  \textbf{(a)} An illustration of the idea of the universe of features.
  Each feature in each image corresponds to a 3D point in the scene.
  We can construct cycle consistent embeddings of the features by mapping each one to the one-hot vector of its corresponding 3D point.
  While there can be many features, there are fewer 3D points and thus this corresponds to a low rank factorization of the correspondence matrix.
  Best viewed in color.
  \textbf{(b)} Visualization of the learned embeddings.
  On the left we have the raw outputs, which are difficult to interpret.
  In the center, we rotated the features to best match the ground truth for a more interpretable visualization (see the end of section \ref{sec:cycconsist}).
  On the right, we have the ground truth embeddings, given as indicator vectors for which feature in the world the points correspond to.
  For the optimally rotated embedding we can see that the true embedding structure is recovered (with some noise).
}
\label{fig:1}
\end{figure}

Our goal is to learn optimal features that capture multiple image views by filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
An outline of our approach can be seen in figure \ref{fig:pipeline}.
We do this by training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Graphs $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ have a set of vertices $\mathcal{V}$ and of edges $\mathcal{E} \subseteq \mathcal{V} \times \mathcal{V}$ and implicitly in the subsequent discussions we assume the edges are directed.
For a vertex $v \in \mathcal{V}$ we use $\mathcal{N}^{h}(v)$ to denote the $h$-hop neighbors of $v$, with the superscript left out for 1-hop neighbors.
Similarly $\mathcal{E}(v)$ is used to denote the edges associated with $v$.
To denote the vertices connected to an edge $e \in \mathcal{E}$ we write $e(v_1, v_2)$.
\subsection{Correspondence Graph} \label{sec:corrgraph}
We assume there is an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, with an associated adjacency matrix $\mA$.
The graph is constructed from putative correspondences of image features across images, typically constructed using feature descriptor distance (e.g. SIFT feature distance).
While there are many interesting methods for computing these putative correspondences \citep{suh2015subgraph, yi2018learning}, we do not explore them in this work.
Typically putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
The ambiguity in the matches could come from repeated structures in the scene, insufficiently informative low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

@@

\subsection{Graph Neural Networks} \label{sec:gnns}
As input to our method we are given a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with the features described in section \ref{sec:corrgraph} $\vf_v \; \forall v \in \mathcal{V}$ and $\vf_e \; \forall e \in \mathcal{E}$.
As with any neural network, GNNs have layered outputs.
We describe the output of layer $k$ as $\vf_v^{(k)} \in \sR^{m_k} \, \forall v \in \mathcal{V}$ and $\vf_e^{(k)} \in \sR^{p_k}\, \forall e \in \mathcal{E}$, with the initial embeddings denoted $\vf_v^{(0)} = \vf_v$ and $\vf_e^{(0)} = \vf_e$.
To aid future analysis, we will represent the features as matrices, denoting the vertex embedding matrix as $\mF_V^{(k)}$ and the edge embedding matrix as $\mF_E^{(k)}$.
If a superscript is not specified then it refers to the final output of the network.

First we describe older methods of GNNs to give context, then we describe the method we use in this work.
Many older methods assume we have the adjacency matrix $\mA$ of the graph known a-priori \cite{bruna2013spectral}, and can encode graph convolutions using the eigenvectors of $\mA$ (these are known as spectral methods).
However, we do not have this luxury, as the correspondence structure changes from image set to image set, and thus we use non-spectral Graph Neural Networks.
Newer models use non-spectral methods, which often ultimately amount to transforming each node with learned weights then averaging each node's representation with its neighbors, known as a message pass \citep{kipf2017semi, defferrard2016convolutional, gama2018mimo, gama2018convolutional}.
Some works such as \cite{gama2019convolutional} use pooling operations on the vertices to make the graph smaller and thus aid computation, but as we need labels on every vertex of the original graph, so we cannot use this.
Most GNNs used in these works can be expressed mathematically as:
\begin{align*}
\tilde{\vf}_v^{(k+1)} =&\; \sigma\left( b^{(k)} + \mW_0^{k} \vf_{v}^{(k)} + \sum_{h=0}^H \sum_{v' \in \mathcal{N}^{h}(v)} f_{e(v,v')} \mW_{h}^{k} \vf_{v'}^{(k)} \right) \\
\end{align*}
The weights/biases $\mW_{h}^{k}$, $b^{(k)}$ are all learned, with no learning done on the edge weights $f_{e(v,v')}$. 
Note that this is just sums or averages over $h$-hop neighborhoods, where the weights on the edges remain static through the computation.
Given that we are trying to prune edges, it would be sensible to add features over edges to learn which ones to prune and which to keep such as in \cite{scarselli2009graph}.

In this work we use the method and implementation described in \cite{battaglia2018relational} (more specifically repetitions of the architecture described in \cite{battaglia2016interaction}).
\cite{battaglia2018relational} uses message passes between node features as well as edge features, which the model can use to prune unnecessary or erroneous edges.
Therefore there is intermediate processing on the edges before information is passed to the vertices.

Mathematically, this is expressed as: % TODO: Maybe make this MLP oriented?
\begin{align}
\tilde{\vf}_{e(v_1,v_2)}^{(k+1)} =&\; \sigma \left(a^{(k)} + \mU_0^{(k)} \vf_{e}^{(k)} + \mU_1^{(k)} \vf_{v_1}^{(k)} + \mU_2^{(k)} \vf_{v_2}^{(k)} \right) \\
\tilde{\vf}_{v}^{(k+1)} =&\; \sigma\left(b^{(k)} + \mW_0^{(k)} \vf_{v}^{(k)} + \sum_{e \in \mathcal{E}(v)} \mW_1^{(k)} \vf_{e}^{(k+1)} \right) 
\end{align}
Here the learned weights are denoted $\mW$ and $\mU$, and the biases $a^{(k)}$ and  $b^{(k)}$.
In \cite{battaglia2018relational}, they allow for more sophisticated aggregation functions, but in this work we simply use the mean function.
Each one of these steps we refer to here as a message pass, and it is analogously equivalent to an iteration in a distributed graph based optimization method.
Between each of the message passes, we further process the features using MLPs.
% In practice, we use MLPs in message passes between vertices and edges for better expressiveness

\subsection{Cycle Consistency} \label{sec:cycconsist}

Let $\mM$ be the noiseless set of matches between our features, with $\mM_{ij}$ being the partial permutation representing the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then for all $i, j, k$:
\begin{equation}
\mM_{ij} = \mM_{ik} \mM_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a ground-truth matching matrix $\mX_i$.
Then the cycle consistency constraint becomes:
\begin{equation}
\mM_{ij} = \mX_{i}\mX_{j}^\top
\label{eq:cycconsist2}
\end{equation}
